The \vspec{} Collection: A suite of utilities to model spectroscopic phase curves of 3D exoplanet atmospheres in the presence of stellar variability.

### 2024-08-05 src/tex/ms.tex
change pypsg name

\newcommand{\urldocsvspecvsm}{\url{https://vspec-vsm.readthedocs.io/en/latest/}}
\newcommand{\urldocspypsg}{\url{https://tedjohnson12.github.io/pypsg/}}
\newcommand{\urldocsgridpolator}{\url{https://vspec-collab.github.io/GridPolator/}}

@@

% Abstract with filler text
\begin{abstract}
    We present the Variable Star PhasE Curve (VSPEC) Collection, a set of Python packages for simulating combined-light spectroscopic observations of 3-dimensional exoplanet atmospheres in the presence of stellar variability and inhomogeneity. \vspec{} uses the Planetary Spectrum Generator's Global Emission Spectra (PSG/GlobES) application along with a custom-built multi-component time-variable stellar model based on an user-defined grid of stellar photosphere models to produce spectroscopic light curves of the planet-host system. \vspec{} can be a useful tool for modeling observations of exoplanets in transiting geometries (primary transit, secondary eclipse) as well as orbital phase curve measurements, and is built in a modular and flexible configuration for easy adaptability to new stellar and planetary model inputs. We additionally present a set of codes developed alongside the core \vspec{} modules, including the stellar surface model generator \texttt{vspec-vsm}, the stellar spectral grid interpolation code GridPolator, and a Python interface for PSG \texttt{pypsg}.
\end{abstract}

@@

To adequately prepare for future observations and future missions, it will be necessary to demonstrate methods to mitigate these effects. This task requires a flexible tool that combines models of exoplanet atmospheres and stellar variability in a robust way. In this paper we present the \vspec{} (Variable Star PhasE Curve\footnote{\url{https://github.com/orgs/VSPEC-collab/repositories}}) Collection, a collection of open-source Python 3 packages to simulate observations of exoplanet systems with variable host stars. The \vspec{}\footnote{\url{https://github.com/VSPEC-collab/VSPEC}} package itself is merely an interface that pulls together a variable star surface model \texttt{vspec-vsm}\footnote{\url{https://github.com/VSPEC-collab/vspec-vsm}}, a stellar spectra grid interpolation code Gridpolator\footnote{\url{https://github.com/VSPEC-collab/GridPolator}}, and a Python interface for the Planetary Spectrum Generator \citep[PSG,][]{villanueva2018}, \texttt{pypsg}\footnote{\url{https://github.com/tedjohnson12/pypsg}}. All of these codes together allow \vspec{} to account for 3D planetary atmospheres, time-resolved effects of both planet and star, multiple geometries including transit and secondary eclipse, and realistic noise modeling.

In this paper we will describe the science motivation behind all of these codes, the details of each package and module in the \vspec{} collection, and demonstrate examples of their use. Section \ref{sec:vspec} describes \vspec{} and goes into some detail about setup, use, and the code's internals. In Section \ref{sec:pypsg} we describe \texttt{pypsg} as an interface for the Planetary Spectrum Generator. Section \ref{sec:star} describes \vspec{}'s stellar model, \texttt{vspec-vsm}. We discuss the spectral model interpolation code GridPolator in Section \ref{sec:gridpolator}. Section \ref{sec:vspec-gcm} is dedicated to the simple climate model that \vspec{} uses to produce realistic phase curves in the absence of a user-supplied output file from a more sophisticated Global Circulation Model (GCM). Section \ref{sec:examples} demonstrates a few simple \vspec{} use cases. Finally, in Section \ref{sec:conclusion} we will discuss the future of the code and issues it might have.

@@

\subsection{Planetary Phase Curve}
\label{subsec:phase-curve}
\vspec{} generates a phase curve though a series of API calls to the Planetary Spectrum Generator \citep[PSG][]{villanueva2018}, using
\texttt{pypsg} (see Section \ref{sec:pypsg}) to interface with the API. Initially, configurations are sent that give PSG information about
the system that does not change with time -- that includes the 3D Global Circulation Model (GCM) and instrument parameters. \vspec{} then enters
the main loop, iterating through each observation epoch (e.g.integration or combination of integrations) and making a pair of API calls: one to get a combined light spectrum and one to get the thermal flux only.
Results from PSG in each epoch are stored locally in the \texttt{.fits} format.

@@

\section{Interfacing with the Planetary Spectrum Generator via \texttt{pypsg}}
\label{sec:pypsg}

@@

\vspec{} uses a Python interface for the PSG API system called \texttt{pypsg}, which is a general stand-alone library built for any Python-based API call to PSG. \texttt{pypsg} draws inspiration from object-relational mapping in frameworks such as Django\footnote{\url{https://www.djangoproject.com/}} to encode a PSG configuration file as a native Python object. Fields of a \texttt{pypsg} data model represent one or more lines of a configuration file, and the two representations are interchangable without loss of information. Upon creation of this configuration as a \texttt{PyConfig} object, a user can create an \texttt{APICall} instance which handles sending a request to PSG via Python's \texttt{requests}\footnote{\url{https://github.com/psf/requests}} package. This dedicated caller is important in that it reads a user's API key from a known location on disk, removing the possibility that the key could be committed to a public repository.

A powerful feature of \texttt{pypsg} is that it provides a universal interface for 3D atmosphere models and PSG's Global Emission Spectra (GlobES) application.
GlobES expects 3D data to be presented in a binary data format, and converting a model to this format is non-trivial. \texttt{pypsg}'s \texttt{PyGCM} class provides a native Python representation of GlobES input, and includes built-in methods for converting popular climate models like ExoCAM \citep{wolf2022} and the Whole Atmosphere Community Climate Model \citep[WACCM,][]{marsh2013} to the correct format.

Listing \ref{ls:pypsg} shows a minimal example that uses \texttt{pypsg} to make a call to PSG/GlobES. Here \texttt{pypsg} converts a GCM output from the netCDF format to a \texttt{PyGCM}, which, along with data model classes like \texttt{Target} and \texttt{Observatory}, is used as an argument to construct a \texttt{PyConfig} class instance. This configuration object is then passed to an \texttt{APICall} instance. When the API call is made, the contents of the \texttt{PyConfig} are included in the body of the POST request in the \texttt{file} field. PSG reads this the same as it would if we had written those contents to a file \texttt{config.txt} and passed the \texttt{--data-urlencode file@config.txt} option to \texttt{cURL}, as are the instructions in the PSG handbook \citep[pp. 162-163]{villanueva2022}. More details on use of this library including its API and examples can be found on the \texttt{pypsg} documentation page\footnote{\urldocspypsg}.

\begin{lstlisting}[
    basicstyle=\footnotesize,language=Python,label={ls:pypsg},caption={\texttt{pypsg} Example}
]
    import pypsg
    from astropy import units as u
    from netCDF4 import Dataset

    # This next line only needs to be run once per user.
    # These settings are saved to `~/.pypsg/settings.json`
    pypsg.settings.save_setting(
        api_key='my_api_key',
        url='https://psg.gsfc.nasa.gov/api.php'
    )

@@

    cfg = pypsg.cfg.PyConfig(
        target=pypsg.cfg.Target(
            object='Exoplanet',
            name='proxima-cen-b'
            ...
        ),
        geometry=pypsg.cfg.Observatory(...),
        generator=pypsg.cfg.Generator(...),
        telescope=pypsg.cfg.SingleTelescope(...)
        noise=pypsg.cfg.CCD(...)
        gcm=pypsg.globes.waccm_to_pygcm(
            gcm_data,
            itime=0,
            molecules=['H2O','CO2'],
            aerosols=['Water','WaterIce'],
            mean_molecular_mass=48.0
        )

@@

    caller = pypsg.APICall(
        cfg=cfg,
        app='globes',
        output_type='rad'
    )
    response = caller()

@@

\section{Simplified Climate Model}
\label{sec:vspec-gcm}
\vspec{} allows users to run simulations with any climate model they desire (so long as it can be converted into the \texttt{pypsg.PyGCM} format) and even comes with utilities to work with PSG/GlobES binary files, ExoCAM \citep{wolf2022}, WACCM \citep{marsh2013}, and ExoPlaSim \citep{paradise2022} without any custom code from the user.

@@

While development is ongoing, these codes are all stable. Future work on \vspec{} will concern improving performance so that it can be used to produce phase-curve grids for retrievals. A goal of \texttt{vspec-vsm} development should be to create a common API between it and more physics-based stellar models so that \vspec{} can interface with those models as well. GridPolator is currently in the process of adding support for more built-in stellar grids, these more complex and covering a larger range of spectral types that the current \vspec{} grid. \texttt{pypsg} has the most to offer the exoplanet community as a stand-alone package, but currently lacks support for much of the data that PSG is capable of producing. Its development so far has focused around the \vspec{} use case, so future work will center around making it more useful to the majority of PSG users.

>>>>

\newcommand{\urldocsvspecvsm}{\url{https://vspec-vsm.readthedocs.io/en/latest/}}
\newcommand{\urldocslibpypsg}{\url{https://tedjohnson12.github.io/libpypsg/}}
\newcommand{\urldocsgridpolator}{\url{https://vspec-collab.github.io/GridPolator/}}

@@

% Abstract with filler text
\begin{abstract}
    We present the Variable Star PhasE Curve (VSPEC) Collection, a set of Python packages for simulating combined-light spectroscopic observations of 3-dimensional exoplanet atmospheres in the presence of stellar variability and inhomogeneity. \vspec{} uses the Planetary Spectrum Generator's Global Emission Spectra (PSG/GlobES) application along with a custom-built multi-component time-variable stellar model based on an user-defined grid of stellar photosphere models to produce spectroscopic light curves of the planet-host system. \vspec{} can be a useful tool for modeling observations of exoplanets in transiting geometries (primary transit, secondary eclipse) as well as orbital phase curve measurements, and is built in a modular and flexible configuration for easy adaptability to new stellar and planetary model inputs. We additionally present a set of codes developed alongside the core \vspec{} modules, including the stellar surface model generator \texttt{vspec-vsm}, the stellar spectral grid interpolation code GridPolator, and a Python interface for PSG \texttt{libpypsg}.
\end{abstract}

@@

To adequately prepare for future observations and future missions, it will be necessary to demonstrate methods to mitigate these effects. This task requires a flexible tool that combines models of exoplanet atmospheres and stellar variability in a robust way. In this paper we present the \vspec{} (Variable Star PhasE Curve\footnote{\url{https://github.com/orgs/VSPEC-collab/repositories}}) Collection, a collection of open-source Python 3 packages to simulate observations of exoplanet systems with variable host stars. The \vspec{}\footnote{\url{https://github.com/VSPEC-collab/VSPEC}} package itself is merely an interface that pulls together a variable star surface model \texttt{vspec-vsm}\footnote{\url{https://github.com/VSPEC-collab/vspec-vsm}}, a stellar spectra grid interpolation code Gridpolator\footnote{\url{https://github.com/VSPEC-collab/GridPolator}}, and a Python interface for the Planetary Spectrum Generator \citep[PSG,][]{villanueva2018}, \texttt{libpypsg}\footnote{\url{https://github.com/tedjohnson12/libpypsg}}. All of these codes together allow \vspec{} to account for 3D planetary atmospheres, time-resolved effects of both planet and star, multiple geometries including transit and secondary eclipse, and realistic noise modeling.

In this paper we will describe the science motivation behind all of these codes, the details of each package and module in the \vspec{} collection, and demonstrate examples of their use. Section \ref{sec:vspec} describes \vspec{} and goes into some detail about setup, use, and the code's internals. In Section \ref{sec:libpypsg} we describe \texttt{libpypsg} as an interface for the Planetary Spectrum Generator. Section \ref{sec:star} describes \vspec{}'s stellar model, \texttt{vspec-vsm}. We discuss the spectral model interpolation code GridPolator in Section \ref{sec:gridpolator}. Section \ref{sec:vspec-gcm} is dedicated to the simple climate model that \vspec{} uses to produce realistic phase curves in the absence of a user-supplied output file from a more sophisticated Global Circulation Model (GCM). Section \ref{sec:examples} demonstrates a few simple \vspec{} use cases. Finally, in Section \ref{sec:conclusion} we will discuss the future of the code and issues it might have.

@@

\subsection{Planetary Phase Curve}
\label{subsec:phase-curve}
\vspec{} generates a phase curve though a series of API calls to the Planetary Spectrum Generator \citep[PSG][]{villanueva2018}, using
\texttt{libpypsg} (see Section \ref{sec:libpypsg}) to interface with the API. Initially, configurations are sent that give PSG information about
the system that does not change with time -- that includes the 3D Global Circulation Model (GCM) and instrument parameters. \vspec{} then enters
the main loop, iterating through each observation epoch (e.g.integration or combination of integrations) and making a pair of API calls: one to get a combined light spectrum and one to get the thermal flux only.
Results from PSG in each epoch are stored locally in the \texttt{.fits} format.

@@

\section{Interfacing with the Planetary Spectrum Generator via \texttt{libpypsg}}
\label{sec:libpypsg}

@@

\vspec{} uses a Python interface for the PSG API system called \texttt{libpypsg}, which is a general stand-alone library built for any Python-based API call to PSG. \texttt{libpypsg} draws inspiration from object-relational mapping in frameworks such as Django\footnote{\url{https://www.djangoproject.com/}} to encode a PSG configuration file as a native Python object. Fields of a \texttt{libpypsg} data model represent one or more lines of a configuration file, and the two representations are interchangable without loss of information. Upon creation of this configuration as a \texttt{PyConfig} object, a user can create an \texttt{APICall} instance which handles sending a request to PSG via Python's \texttt{requests}\footnote{\url{https://github.com/psf/requests}} package. This dedicated caller is important in that it reads a user's API key from a known location on disk, removing the possibility that the key could be committed to a public repository.

A powerful feature of \texttt{libpypsg} is that it provides a universal interface for 3D atmosphere models and PSG's Global Emission Spectra (GlobES) application.
GlobES expects 3D data to be presented in a binary data format, and converting a model to this format is non-trivial. \texttt{libpypsg}'s \texttt{PyGCM} class provides a native Python representation of GlobES input, and includes built-in methods for converting popular climate models like ExoCAM \citep{wolf2022} and the Whole Atmosphere Community Climate Model \citep[WACCM,][]{marsh2013} to the correct format.

Listing \ref{ls:libpypsg} shows a minimal example that uses \texttt{libpypsg} to make a call to PSG/GlobES. Here \texttt{libpypsg} converts a GCM output from the netCDF format to a \texttt{PyGCM}, which, along with data model classes like \texttt{Target} and \texttt{Observatory}, is used as an argument to construct a \texttt{PyConfig} class instance. This configuration object is then passed to an \texttt{APICall} instance. When the API call is made, the contents of the \texttt{PyConfig} are included in the body of the POST request in the \texttt{file} field. PSG reads this the same as it would if we had written those contents to a file \texttt{config.txt} and passed the \texttt{--data-urlencode file@config.txt} option to \texttt{cURL}, as are the instructions in the PSG handbook \citep[pp. 162-163]{villanueva2022}. More details on use of this library including its API and examples can be found on the \texttt{libpypsg} documentation page\footnote{\urldocslibpypsg}.

\begin{lstlisting}[
    basicstyle=\footnotesize,language=Python,label={ls:libpypsg},caption={\texttt{libpypsg} Example}
]
    import libpypsg
    from astropy import units as u
    from netCDF4 import Dataset

    # This next line only needs to be run once per user.
    #     These settings are 
    #     saved to `~/.libpypsg/settings.json`
    libpypsg.settings.save_setting(
        api_key='my_api_key',
        url='https://psg.gsfc.nasa.gov/api.php'
    )

@@

    cfg = libpypsg.cfg.PyConfig(
        target=libpypsg.cfg.Target(
            object='Exoplanet',
            name='proxima-cen-b'
            ...
        ),
        geometry=libpypsg.cfg.Observatory(...),
        generator=libpypsg.cfg.Generator(...),
        telescope=libpypsg.cfg.SingleTelescope(...)
        noise=libpypsg.cfg.CCD(...)
        gcm=libpypsg.globes.waccm_to_pygcm(
            gcm_data,
            itime=0,
            molecules=['H2O','CO2'],
            aerosols=['Water','WaterIce'],
            mean_molecular_mass=48.0
        )

@@

    caller = libpypsg.APICall(
        cfg=cfg,
        app='globes',
        output_type='rad'
    )
    response = caller()

@@

\section{Simplified Climate Model}
\label{sec:vspec-gcm}
\vspec{} allows users to run simulations with any climate model they desire (so long as it can be converted into the \texttt{libpypsg.PyGCM} format) and even comes with utilities to work with PSG/GlobES binary files, ExoCAM \citep{wolf2022}, WACCM \citep{marsh2013}, and ExoPlaSim \citep{paradise2022} without any custom code from the user.

@@

While development is ongoing, these codes are all stable. Future work on \vspec{} will concern improving performance so that it can be used to produce phase-curve grids for retrievals. A goal of \texttt{vspec-vsm} development should be to create a common API between it and more physics-based stellar models so that \vspec{} can interface with those models as well. GridPolator is currently in the process of adding support for more built-in stellar grids, these more complex and covering a larger range of spectral types that the current \vspec{} grid. \texttt{libpypsg} has the most to offer the exoplanet community as a stand-alone package, but currently lacks support for much of the data that PSG is capable of producing. Its development so far has focused around the \vspec{} use case, so future work will center around making it more useful to the majority of PSG users.
